### 2015-05-24 implementation.tex
patch openssh, implem openssh and drivers.

\lstset{language=c}
\lstset{commentstyle=\color{colComments}\textit,
float=hbp,%
basicstyle=\ttfamily\footnotesize, %
identifierstyle=\color{colIdentifier}, %
keywordstyle=\color{colKeys}, %
stringstyle=\color{colString}, %
columns=flexible, %
tabsize=2, %
extendedchars=true, %
showspaces=false, %
showstringspaces=false, %
numbers=left, %
numberstyle=\tiny, %
breaklines=true, %
breakautoindent=true, %
captionpos=b}
\begin{lstlisting}[caption=openvpn compress then encrypt -- sample from \texttt{forward.c}, label=list:openvpn-workflow]
/* Compress, fragment, encrypt and HMAC-sign an outgoing packet. */
void encrypt_sign (struct context *c, bool comp_frag)
{
	struct context_buffers *b = c->c2.buffers;
	const uint8_t *orig_buf = c->c2.buf.data;

@@

\subsection{OpenSSH}
% Talk about the MAC=none patch
% connection using SSL/TLS, then AES, but without fragmentation (unlike openvpn)

@@

It is to be noted that the present work uses an implementation with all the debug flags activated.
If it does have an impact on the performance, it is minimal: in the worst case of the benchmark, the throughput drops only by 2.4\%.
Moreover, the benchmark maximizes this difference by doing only OpenSSL operations.
When OpenSSL will have to share the CPU with other applications, the loss will be even less noticeable.
\begin{figure}[ht]
\begin{tikzpicture}
\begin{axis}[
		title = {Openssl speed with debug on/off},
        width  = 0.85*\textwidth,
        height = 8cm,
        major x tick style = transparent,
        ybar,
        bar width=8pt,
        ymajorgrids = true,
        ylabel = {Throughput [$10^3$kB/s]},
        xlabel = {Packet size},
        ymin=0,
        symbolic x coords={16,64,256,1024,8192},
        xtick = data,
        scaled y ticks = false,%Disable the *10^4 exponent applied to all y axis markings.
        legend style={at={(0.5,-0.25)},	anchor=north,legend columns=2},
        enlarge x limits=0.1,
    ]
\addplot[style={MidnightBlue,fill=MidnightBlue,mark=none}]
	coordinates {(16,18400.60) (64,20750.61) (256,21507.33) (1024,21681.15) (8192,21725.18)};
	\label{aes-256-cbc-dbg-off}

@@

\legend{AES-256-CBC debug [off], AES-256-CBC debug [on], AES-128-CBC debug [off], AES-128-CBC debug [on]}
\end{axis}
\end{tikzpicture}
\label{fig:openssl-speed-dbg-on-off}
\caption{Software benchmark of Openssl speed for AES mode CBC, with 128- and 256-bit keys, debugging flags (de)activated at compilation (\texttt{-fno-inline -g -marm}).}{}
\end{figure}

@@

\subsection{Linux drivers}
% Mention the fact that we use the assembly implementation of AES, not the generic C one.

@@

@@

\subsection{BA414E Driver and Silex engine}
At the moment of development, there is no asymetric cryptography interface in the crypto API.
The BA414E can thus not be accessed using the same path as the BA411E and a user-space driver will be needed, that is the leftern path of figure~\ref{fig:os-path-generic-b}.
As presented in section~\ref{sec:theory-driver}, we will need to rely on an UIO driver to access the device from the user space.

>>>>

\lstset{language=c}
\begin{lstlisting}[caption=openvpn compress then encrypt -- sample from \texttt{forward.c}, label=list:openvpn-workflow]
/* Compress, fragment, encrypt and HMAC-sign an outgoing packet. */
void encrypt_sign (struct context *c, bool comp_frag)
{
	struct context_buffers *b = c->c2.buffers;
	const uint8_t *orig_buf = c->c2.buf.data;

@@

\subsection{OpenSSH}
% Talk about the MAC=none patch
% connection using SSL/TLS, then AES, but without fragmentation (unlike openvpn)
OpenSSH relies on an external cryptographic library for all its security operations.
Up until the version 6.7 published in october 2014, it had to be compiled against OpenSSL.
However, after the infamous security vulnerability heartbleed in april 2014, the developpers took a step to move towards LibreSSL, a fork of OpenSSL managed by OpenBSD  developers.
Still, there is no official support for any other cryptographic library.

If OpenSSH does support most of OpenSSL ciphers by default, it takes some liberties such as disabling the CBC encryption mode and removing the support of no MAC during a transmission.

\noindent The first liberty is a laste\footnote{A vulnerability note as been issued by Carnegie Mellon University Computer Emergency Response Team in early 2009 (last revision)~\cite{CERT2009}, in response to a research of University of London~\cite{Albrecht:2009} presenting a plaintext-recovering attack against SSH when CBC mode is used, but the OpenSSH update took only place in october 2014.}.
Even if it is not recommended, the user can still enable this mode by explicitly configuring it in the \texttt{sshd\_config} options.

\noindent The second liberty has been taken to prevent the user to strip himself from data authenticity and integrity.
However, in the case of testing and benchmarking, leaving the MAC behind can be interresting, especially if it is not offloaded in hardware such as the encryption in our case.
In order to add this feature to OpenSSH, we wrote a patch to apply on OpenSSH 6.7, available in appendix~\cite{chap:openssh-patch}.

@@

It is to be noted that the present work uses an implementation with all the debug flags activated.
Figure~\ref{fig:openssl-speed-dbg-on-off} shows that if it does have an impact on the performance, it is minimal: in the worst case of the benchmark, the throughput drops only by 2.4\%.
Moreover, the benchmark maximizes this difference by doing only OpenSSL operations.
When OpenSSL will have to share the CPU with other applications, the loss will be even less noticeable.
\begin{figure}[ht]
\begin{tikzpicture}
\begin{axis}[
		title = {Openssl speed with debug on/off},
        width  = 0.85*\textwidth,
        height = 8cm,
        major x tick style = transparent,
        ybar,
        bar width=8pt,
        ymajorgrids = true,
        ylabel = {Throughput [$10^3$kB/s]},
        xlabel = {Packet size},
        ymin=0,
        symbolic x coords={16,64,256,1024,8192},
        xtick = data,
        scaled y ticks = false,%Disable the *10^4 exponent applied to all y axis markings.
        legend style={at={(0.5,-0.25)},	anchor=north,legend columns=2},
        enlarge x limits=0.1,
    ]
\addplot[style={MidnightBlue,fill=MidnightBlue,mark=none}]
	coordinates {(16,18400.60) (64,20750.61) (256,21507.33) (1024,21681.15) (8192,21725.18)};
	\label{aes-256-cbc-dbg-off}

@@

\legend{AES-256-CBC debug [off], AES-256-CBC debug [on], AES-128-CBC debug [off], AES-128-CBC debug [on]}
\end{axis}
\end{tikzpicture}
\caption{OpenSSL debugging benchmark}{Software benchmark of Openssl speed for AES mode CBC, with 128- and 256-bit keys, debugging flags (de)activated at compilation (\texttt{-fno-inline -g -marm}). The throughput difference ranges from 0.2\% and 2.4\% , and is more marked for larger keysize, as more debugging data needs to be generated.}
\label{fig:openssl-speed-dbg-on-off}
\end{figure}

@@

\subsection{Linux drivers}
% Mention the fact that we use the assembly implementation of AES, not the generic C one.
% Need an lsmod here.
Several kernel modules are needed to implement the various cryptographic algorithm in software.
The GCM alone needs five different modules, and IPSec three others.
The following description addresses all the kernel modules required to run all the use cases presented in this work.

\begin{description}
	\item[\texttt{aes\_arm}] Assembly implementation of AES. This version is optimized to use the ARMv7 instruction set.
	\item[\texttt{sha256-generic}] C implementation of SHA-256.
	\item[\texttt{gfmult128}] Multiplication in $GF(2^128)$, needed by the GCM mode.
	\item[\texttt{ghash}] GHASH function needed by the authentication of the GCM mode.
	\item[\texttt{seqiv}] Sequence IV number generator, needed by the CTR and GCM modes.
	\item[\texttt{cbc}] C based CBC mode.
	\item[\texttt{gcm}] C base GCM mode.
	\item[\texttt{crypto\_null}] Null cipher. This basically does nothing on the plaintext, but it is still needed to propose the interface in the kernel.
	\item[\texttt{tun}] TUN/TAP network device, needed by OpenVPN.
	\item[\texttt{xfrm\_user}] XFRM operations.
	\item[\texttt{esp4}] IPv4 ESP implementation. The AH counterparty is also available in the \texttt{ah4} module.
	\item[\texttt{ipcomp}] IP compression module. Needed by IPSec if the option is activated in the configuration.
	\item[\texttt{cryptodev}] Creates \texttt{/dev/crypto}, giving access to the crypto API in the user space.
	\item[\texttt{uio}] User-space I/O driver, allowing to access the hardware memory from the user space, needed by our implementation of the BA414E driver.
\end{description}

@@

The table~\ref{tab:ip-ciphers} summarizes the ciphers supported by the two Barco Silex' IPs used in this work.

@@

\subsection{BA414E Driver and Silex engine}
At the moment of development, there is no asymetric cryptography interface in the crypto API\footnote{A request for comment patch as been submitted to the Linux kernel mailing list~\cite{crypto-api-pk-encryption} in late April 2015, proposing a standard interface for public key encryption in the crypto API.}.
The BA414E can thus not be accessed using the same path as the BA411E and a user-space driver will be needed, that is the leftern path of figure~\ref{fig:os-path-generic-b}.
As presented in section~\ref{sec:theory-driver}, we will need to rely on an UIO driver to access the device from the user space.
